Open archaeology, open source?

### 2024-01-24 analysis/_03-data_methodology.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

```{r data-oarch}
read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") |>
  mutate(category = na_if(category, "?")) ->
  oarch

@@

```{r tbl-glossary}
#| tbl-cap: Glossary of git and GitHub terminology
tribble(
  ~term, ~definition,
  "Comment", "On GitHub, text post attached to an issue, including the first one that describes the issue",
  "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
  "Commit access", "Ability to make changes to a repository directly, without making a pull request",
  "Contributor", "User that has made at least one commit to a specified repository",
  "Follow", "Add activity by another user to a user's timeline",
  "Forge", "Web-based platform for hosting, distributing and facilitating collaboration on version controlled computer code, e.g. GitHub, GitLab, Codeberg",
  "Fork", "Copy of a repository owned by another user; forking is a prerequisite to making a pull request",
  "git", "Open source version control software",
  "GitHub", "Commercial platform that freely hosts git repositories and provides extended collaboration and social networking features, such as pull requests, issues and stars",
  "GitLab", "Open source alternative to GitHub",
  "CodeBerg", "Open source alternative to GitHub",
  "Issue", "Feature of GitHub that records and tracks a bug report, feature request or other suggestion in a repository",
  "Maintainer", "Individual that has overall control of a repository, generally assumed to be its primary contributor. Repositories can have multiple users with commit access in addition to the maintainer.",
  "Merge", "Accept a pull request and incorporate its changes into a repository.",
  "Organization", "Entity representing a group of users, which can also own repositories",
  "Pull request", "Mechanism by which users that don't have commit access to a repository can contribute to it. The repository's maintainer or another user with commit access must decide whether to merge (accept) the changes, or decline them.",
  "Repository", "Individual project that uses git for version control. Can include a mix of different types of files.",
  "Star", "GitHub's version of a 'like', applied by users to a repository.",
  "Timeline", "Chronological feed of GitHub activity from repositories a user has starred and other users they follow. Also includes repositories that a user is not following if they are 'trending' or determined relevant by GitHub's algorithm.",
  "User", "On GitHub, an individual with an account that can own repositories",
  "Version control", "System for tracking changes (additions, alterations, or deletions) in a set of files, typically but not exclusively computer code"
) |>
  gt() |>
  cols_label_with(fn = tools::toTitleCase) |>
  fmt_markdown(definition)
```

@@

```{r tbl-forges}
#| tbl-cap: Software forges used by open-archaeo projects
gt(oarch_forges) |>
  cols_label(repo_host = "Host", total = "Total") |>
  cols_merge_n_pct(total, p_total) |>
  fmt_percent(p_total) |>
  tab_spanner("By category", Guides:`Standalone software`)
```

>>>>

```{r data-oarch}
# TODO: update and replace with local file before submission
read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") |>
  mutate(category = na_if(category, "?")) ->
  oarch

@@

```{r tbl-glossary}
#| tbl-cap: Glossary of git and GitHub terminology
tribble(
  ~term, ~definition,
  "Comment", "On GitHub, text post attached to an issue, including the first one that describes the issue",
  "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
  "Commit access", "Ability to make changes to a repository directly, without making a pull request",
  "Contributor", "User that has made at least one commit to a specified repository",
  "Follow", "Add activity by another user to a user's timeline",
  "Forge", "Web-based platform for hosting, distributing and facilitating collaboration on version controlled computer code, e.g. GitHub, GitLab, Codeberg",
  "Fork", "Copy of a repository owned by another user; forking is a prerequisite to making a pull request",
  "git", "Open source version control software",
  "GitHub", "Commercial platform that freely hosts git repositories and provides extended collaboration and social networking features, such as pull requests, issues and stars",
  "GitLab", "Open source alternative to GitHub",
  "CodeBerg", "Open source alternative to GitHub",
  "Issue", "Feature of GitHub that records and tracks a bug report, feature request or other suggestion in a repository",
  "Maintainer", "Individual that has overall control of a repository, generally assumed to be its primary contributor. Repositories can have multiple users with commit access in addition to the maintainer.",
  "Merge", "Accept a pull request and incorporate its changes into a repository.",
  "Organization", "Entity representing a group of users, which can also own repositories",
  "Pull request", "Mechanism by which users that don't have commit access to a repository can contribute to it. The repository's maintainer or another user with commit access must decide whether to merge (accept) the changes, or decline them.",
  "Repository", "Individual project that uses git for version control. Can include a mix of different types of files.",
  "Star", "GitHub's version of a 'like', applied by users to a repository.",
  "Timeline", "Chronological feed of GitHub activity from repositories a user has starred and other users they follow. Also includes repositories that a user is not following if they are 'trending' or determined relevant by GitHub's algorithm.",
  "User", "On GitHub, an individual with an account that can own repositories",
  "Version control", "System for tracking changes (additions, alterations, or deletions) in a set of files, typically but not exclusively computer code"
) |>
  rename_with(tools::toTitleCase) |>
  tt() |>
  format_tt(2, markdown = TRUE)
```

@@

```{r tbl-forges}
#| tbl-cap: Software forges used by open-archaeo projects
oarch_forges |>
  mutate(
    total = glue("{total} ({format(p_total * 100, trim = TRUE, digits = 1)}%)")
  ) |>
  select(-p_total) |>
  rename(Host = repo_host, Total = total) |>
  tt()
```

### 2024-01-24 analysis/_04-open_archaeology.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

```{r tbl-categories}
#| tbl-cap: Categories of open archaeology projects included in open-archaeo.
# TODO: Include sum for the "Software" and "Documents" supercategories, and make them appear in bold.
tribble(
  ~category,                               ~kind,       ~scope,
  "Packages and libraries",                "Software", "Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.",
  "Standalone software",                   "Software", "Software that may be operated without needing to first access an underlying platform.",
  "Scripts",                               "Software", "Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.",
  "Specifications, protocols and schemas", "Documents", "A formal data structure or framework intended to be used as a model.",
  "Products",                              "Documents", "Stable outcomes of creative work.",
  "Guides",                                "Documents", "An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.",
  "Lists and datasets",                    "Documents", "A series of consistently organized observations assembled with purpose."
) |>
  left_join(oarch, by = "category") |>
  count(category, kind, scope) |>
  group_by(kind) |>
  arrange(-n) |>
  gt() |>
  cols_label_with(fn = tools::toTitleCase)
```

@@

```{r tbl-platforms}
#| tbl-cap: Platforms and programming languages used by open archaeology projects.
oarch |>
  drop_na(platform) |>
  count(platform) |>
  mutate(
    significant = n > 1,
    platform = if_else(significant, platform, "Other")
  ) |>
  group_by(platform) |>
  summarise(n = sum(n)) |>
  arrange(platform == "Other", -n) |>
  gt() |>
  cols_label_with(fn = tools::toTitleCase)
```

@@

```{r tbl-repo-licenses}
#| tbl-cap: Summary of licenses used by projects with GitHub repositories
# TODO: group small categories into "other"
oarch |>
  mutate(license = str_remove(license, "-?\\d\\.\\d$")) |>
  count(license) |>
  mutate(license = replace_na(license, "None detected")) |>
  arrange(-n) |>
  gt() |>
    cols_label(license = "License", n = "N")
```

>>>>

```{r tbl-categories}
#| tbl-cap: Categories of open archaeology projects included in open-archaeo.
# TODO: Include sum for the "Software" and "Documents" supercategories, and make them appear in bold.
tribble(
  ~category,                               ~kind,       ~scope,
  "Packages and libraries",                "Software", "Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.",
  "Standalone software",                   "Software", "Software that may be operated without needing to first access an underlying platform.",
  "Scripts",                               "Software", "Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.",
  "Specifications, protocols and schemas", "Documents", "A formal data structure or framework intended to be used as a model.",
  "Products",                              "Documents", "Stable outcomes of creative work.",
  "Guides",                                "Documents", "An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.",
  "Lists and datasets",                    "Documents", "A series of consistently organized observations assembled with purpose."
) |>
  left_join(oarch, by = "category") |>
  count(category, kind, scope) |>
  group_by(kind) |>
  arrange(-n) |>
  rename_with(tools::toTitleCase) |>
  tt()
```

@@

```{r tbl-platforms}
#| tbl-cap: Platforms and programming languages used by open archaeology projects.
oarch |>
  drop_na(platform) |>
  count(platform) |>
  mutate(
    significant = n > 1,
    platform = if_else(significant, platform, "Other")
  ) |>
  group_by(platform) |>
  summarise(n = sum(n)) |>
  arrange(platform == "Other", -n) |>
  rename_with(tools::toTitleCase) |>
  tt()
```

@@

```{r tbl-repo-licenses}
#| tbl-cap: Summary of licenses used by projects with GitHub repositories
# TODO: group small categories into "other"
oarch |>
  mutate(license = str_remove(license, "-?\\d\\.\\d$")) |>
  count(license) |>
  mutate(license = replace_na(license, "None detected")) |>
  arrange(-n) |>
  rename_with(tools::toTitleCase) |>
  tt()
```

### 2024-01-24 analysis/_06-network.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

```{r tbl-repo-centrality}
#| tbl-cap: Most central repositories in the repository–repository network
oarch_graph_repo_repo |>
  mutate(centrality = centrality_betweenness()) |>
  as_tibble() |>
  arrange(-centrality) |>
  slice(1:10) |>
  left_join(oarch, c("repo" = "gh_repo")) |>
  mutate(n_commits = map_int(gh_commits, nrow)) |>
  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
  select(repo, centrality, category, tags, n_commits) |>
  gt() |>
  cols_label(
    repo = "Repository", 
    centrality = "Betweenness Centrality",
    category = "Category",
    tags = "Tags",
    n_commits = "Commits"
  )
```

>>>>

```{r tbl-repo-centrality}
#| tbl-cap: Most central repositories in the repository–repository network
oarch_graph_repo_repo |>
  mutate(centrality = centrality_betweenness()) |>
  as_tibble() |>
  arrange(-centrality) |>
  slice(1:10) |>
  left_join(oarch, c("repo" = "gh_repo")) |>
  mutate(n_commits = map_int(gh_commits, nrow)) |>
  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
  select(repo, centrality, category, tags, n_commits) |>
  rename(repository = repo, commits = n_commits) |>
  rename_with(tools::toTitleCase) |>
  tt()
```

### 2024-01-24 analysis/openarchaeo_collaboration.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

library("english")
library("fs")
library("geomtextpath")
library("ggiraph")
library("ggraph")
library("gt")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("scales")
library("tidygraph")
library("tidyverse")
```

>>>>

library("english")
library("fs")
library("geomtextpath")
library("ggiraph")
library("ggraph")
library("here")
library("khroma")
library("lubridate")
library("patchwork")
library("scales")
library("tidygraph")
library("tidyverse")
library("tinytable")
```
